A Non-local Bayesian Video Denoising Method

### 2015-10-16 doc/201507_tip/paper.tex
[doc] TIP paper text almost finished. Still missing updated experiments.

Advances in video sensor hardware have steadily improved the acquisition
quality. However, due to the reduction in the price of cameras
and data storage, video cameras are being used more each time, and in less
favorable situations, such as low lighting conditions. This results in high 
levels of noise, which negatively affects the visual quality of the video and
hinders its use for any application. 
\todo{Add references for claims about sensors, camera prices, etc.}

Video denoising has received much less attention than still image denoising. 
In principle, video denoising should be easier than image denoising, since the
strong temporal redundancy along motion trajectories favours the denoising
task. However, this additional form of redundancy creates challenges.  On one
hand, the amount of data is much larger than for still images, and efficient
ways of navigating through this data are needed. On the other hand, the output
of the denoising algorithm should be temporally consistent according to the
motion of the original sequence.

@@

Some methods do not distinguish between the temporal and spatial dimensions,
and treat the video as a volume, which is denoised in a transformed domain
\cite{Rajpoot2003,Wilson2004,Selesnick2003}. 
\todo{Read these references. Even if they consider a 3D transform, I think
they DO distinguish between space and time}

Until the advent of patch-based approaches, methods that do not compensate
motion failed for sequences with moderate motion, and their main interest resided
in their simplicity and low computational cost. First state-of-the-art results
obtained without motion estimation were reported in \cite{Buades2005v}, with the
non-local means algorithm. The authors argued that for their method, based on
averaging similar patches, motion estimation was not only unnecessary but
counterproductive: while textured regions are problematic for motion
estimation, they are a source of a large number of similar patches across
different frames. Even if the motion trajectory could be reliably estimated, it
makes more sense to use all similar patches, rather than using only the ones on the
trajectory.
%
A similar approach was followed by the authors of \cite{Dabov2007v}. This method is based
on collaborative filtering of similar patches in a
spatio-temporal neighborhood.

The methods in \cite{Buades2005v} and \cite{Dabov2007v} are extensions to video
of still image patch-based denoising algorithms (non-local means REF and BM3D
REF respectively). They are based on \suggest{filtering} similar 2D patches
searched for in a 3D spatio-temporal neighborhood.  These methods exploit both
spatial and temporal redundancy, however, since each 2D patch is processed
independently, there is no mechanism to impose coherence along trajectories.
This results in flickering artifacts which become noticeable for high levels of
noise. 
%
While these methods exploit spatio-temporal redundancy, they do not impose
coherence of the trajectories, resulting in flickering
artifacts which become noticeable for high levels of noise.

@@

In this work we present an extension to video of the non-local Bayes image
denoising method \cite{Lebrun2013a,Lebrun2013ipol} using 3D rectangular
patches and which does not require motion estimation. Yet, it provides
results which compare favourably to the current state-of-the-art,
both quantitatively by a significant marging and qualitatively, specifically in
terms of the temporal consistency attained.

@@

The rest of the paper is structured as follows: in \S
\ref{sec:review_nonlocal_Bayes} we review the Bayesian patch denoising principle 
and describe our extension to video. In \S \ref{sec:parameters}
we discuss its parameters. Some results, including a comparison with VBM4D, are shown
in \S \ref{sec:results}. Concluding remarks are given in \S \ref{sec:conclusion}

@@

The Figures \ref{fig:patch_groups_pos} and \ref{fig:patch_groups_patches}
show two example of groups of similar patches. Figure \ref{fig:patch_group_pos}
shows the position of 200 nearest neighbors of a reference patch. It can be noticed
that inspite of the noise

@@

\begin{figure*}[htpb!]
	\centering
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}\\
	\vspace{.2cm}
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
%	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
	\caption{Examples of sets of 200 nearest neighbors spatio-temporal patches,
	for patches of size $9\times 9\times 4$.
	The images in each row show the five frames of a $37\times37\times5$ search region. 
	In red we show the positions of the 5 nearest neighbors, in green the next 40 and in blue the rest.
	Note the points shown correspond to the top-left pixel of the first temporal
	slice of each patch.
	The reference patch is the red patch in the center of the search region.}
	\label{fig:patch_groups_pos}
\end{figure*}

\begin{figure}[htpb!]
	\centering
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
	\vspace{.2cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
	\vspace{.2cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
%	\vspace{.1cm}
	\caption{For each of the two sets of patches of Fig. \ref{fig:patch_groups_pos} we show
	the 20 nearest neighbors (top), the corresponding MAP estimates (middle) and the sample mean 
	and first 19 principal directions computed from the sample covariance matrix (bottom).
	For each image, each column corresponds to the 4 temporal slices of a $9\times9\times4$ patch.}
	\label{fig:patch_groups_patches}
\end{figure}

\section{Discussion}

\subsection{Spatio-temporal patches}

Patch-based methods filter similar patches to remove noise. When the noise is
Gaussian, the patch similarity is determined based on the (squared) Euclidean distance 
between patches. Due to the noise, the Euclidean distance follows a non-central
$\chi^2$ distribution, with variance
\[\textnormal{Var}\left\{\frac1{m}\|\ma q_1 - \ma q_2\|^2\right\} = \frac{8\sigma^2}{m}\left(\sigma^2 + \frac1{m}\|\ma p_1 - \ma p_2\|^2\right),\]
where $\ma q_1, \ma q_2$ are the noisy versions of the patches $\ma p_1, \ma p_2$, 
and $m$ denotes the number of pixels in the patch. The noise in the patch comparison
can be reduced by considering larger patches.
However, increasing the size of the patch also increases the distances between
patches and reduces the likelihood of finding similar patches.
%
The additional temporal dimension in a spatio-temporal patch allows to increase
the number of pixels in the patch, without increasing its spatial size.
Due to the high redundancy of the video in the temporal dimension, increasing the 
temporal size of the patch causes a much lower increase in the patch distances.
%The high redundancy of the video in the temporal dimension, allows to have 
%larger patches without affecting much the number of nearest neighbors of the patch.
%

@@

\pcomment{
\begin{itemize}
%	\item Justify why 3D patches: Greater distance noise reduction with roughly the same 
%		degree of self-similarity. One way to see this is that the 3D patches can be 
%		seen, for a video in which the brightness constancy assumption holds, as a 2D patch
%		by back-projecting all the slices to the first frame.
	\item Relation with respect other iterative methods found in the literature
		(SOS, twicing, diffusion, BM3D, I-SA-BM4D).
	\item Variation of the results tables for different sequences.
\end{itemize}}

\subsection{Estimation of the covariance matrix}

\section{Description of the algorithm}

\subsection{Implementation details}

\paragraph{Speed-ups} To accelerate the computation we apply two tricks
considered in \cite{Lebrun2013ipol}. The first one visits a patch each $n$
patches (similar to \cite{Dabov2007tip}). The second trick consists on the following.
After filtering the patches in a group, these patches are not considered again
as reference patches for a group. This reduces the number of groups of patches
that need to be processed by a factor of $n$.
\todo{Also describe that the 4 neighbors are flagged as processed.}

\subsection{Computational complexity}
\label{sec:complexity}

Let us consider the complexity of constructing and filtering a group of
$n$ similar patches. Due to the speed-ups described in the
previous section, the actual number of groups of similar patches processed
might vary between a $1\%$ and a $10\%$ of the total number of pixels in the
video.

Processing each group of similar patches is of asymptotic order $\mathcal
O(w_x^2 w_t s_x^2 s_t + s_x^4s_t^2r + ns_x^4s_t^2)$.
The first term comes from the exhaustive search of the $n$ most
similar patches in the search window. 
The second term comes from the computation of the truncated eigen-decomposition
of the covariance matrix, and the last term contemplates computing the
covariance matrix.

The most expensive operations are the computation of the covariance matrix and
of its principal directions.
Let us note that there are methods for computing the $r$ principal directions
of a set of points that do not require computing the covariance matrix\todo{referencia}. Instead
these methods efficiently approximate a truncated SVD of the data matrix, and would 
allow to improve the computational complexity to $\mathcal O(w_x^2 w_t s_x^2
s_t + s_x^2s_trn)$.

\todo{Statistics of group of patches.}

\subsection{Selection of parameters}
\label{sec:parameters}

@@

\paragraph{Search region} Setting the size of the search region implies a
trade-off between denoising performance and running time: A larger search
region takes longer to traverse but yields better results, since more
similar patches can be found. We fix the search
region for both stages of the algorithm at 
\begin{equation}
w_x = 37 \quad \text{ and } \quad w_t = 5.
\label{eq:search_region_parameters}
\end{equation}
With these values, the search region captures the $5$ frame trajectory of the
target patch as long as it moves slower than $37/5 = 7.4$
pixels/frame.

\paragraph{Patch size} Let us consider first the spatial patch
size $s_x$. This is a key parameter in patch-based methods that
exploit the image self-similarity \pcomment{CITAS}. The distribution of the 
patch distances is a non-central Chi-squared distribution \cite{Dabov2007tip}
\pcomment{CHECK!}
with variance 
\[\textnormal{Var}\left\{\frac1{m}\|\ma q_1 - \ma q_2\|^2\right\} = \frac{8\sigma^2}{m}\left(\sigma^2 + \frac{\|\ma p_1 - \ma p_2\|^2}{m}\right),\]
where $m = s_x^2s_t$ is the total number of pixels in the patch.
Increasing the spatial patch size $s_x$ reduces the variance of the noise in the patch
distance. 
However, for larger
patches it becomes less likely to find other patches with a similar texture
pattern. The use spatio-temporal patches helps in solving this trade-off: for
instance, with a spatio-temporal patch of size $5\times 5\times 4$, we obtain
the same patch distance noise as with a spatial patch of size $10\times 10
\times 1$. Moreover, if the motion in the video is approximately uniform, then the 
amount of similar neighbors of a $5\times 5\times 4$ patch is roughly the same
as for a $5\times 5\times 1$ patch.
\pcomment{This discussion would be nice to have in the introduction. Furthermore,
we could add a short computation on the relation between the distance between
3D patches and their 2D lids. In this way, we could speak of an equivalent 
2D patch size.}

@@

\paragraph{Rank and number of similar patches} These parameters are intimately
linked. Together they determine the estimation of the local Gaussian models
and thus control the bias-variance trade-off of the denoiser.
\pcomment{This is probably inaccurate. There should be other parameters
contributing to this.}

The number of similar patches $n$ affects the estimation of the covariance
matrix $C$.
%Let us consider the trade-offs involved in the estimation of the covariance
%matrix $C$.
The empirical covariance matrix $\widehat C$ converges to the true covariance matrix
$C$ as the number of samples from the underlying Gaussian distribution grows to infinity.
In our case, the number of samples we can ``draw'' from the Gaussian is limited:
%The Gaussian is only an approximate density model for the set of
%$n$ nearest neighbors of a target patch.
As we increase the number $n$ of neighbors of a patch, we include patches
that are not as similar to the reference patch and the Gaussian model becomes less valid. 
%
Thus there is a trade-off between accuracy of Gaussian model and precision of the
covariance matrix.
We find a compromise to this trade-off empirically, by trying different values of this parameter on a 
few training sequences and use the one yielding the best PSNR.

@@

The plots in Figure \ref{fig:rank1-nsim1-bpsnr} show the PSNR of the basic
estimate as a function of the rank $r_1$, for several values of $n_{1}$.
The curves correspond to an average over 20 frames clips 
extracted from four classical test
sequences, namely \emph{bus}, \emph{foreman}, \emph{football} and \emph{tennis}.
The sequences were contaminated with AGWN of standard deviation $\sigma = 10,
20, 30, 40$. 
%
%
We can interpret the results in the light of the theoretical results
described above.
For all noise levels, best results are obtained when using a larger number of
samples $n_1$. All curves of fixed $n_1$ show a maximum at a given rank lower than
the dimensionality of the patch. As the noise level increases, this optimal rank
shifts to the left. This makes sense, since less eigenvectors are reliable estimates.
%
Also, for a fixed noise level, the optimal rank is smaller for smaller values of $n_1$.
Smaller groups obtained by keeping only the
most similar patches have less modes of variation, and thus a worse signal-to-noise ratio
in the eigenvectors with lower eigenvalues.
%As the group of patches grows
%bigger, patches in the group are not as similar anymore, and a higher rank is
%needed to allow for additional modes of variation.

@@

In this section we present results obtained with the proposed
method on some classic test sequences with Gaussian noise added.
We consider separatedly the cases of grayscale and color videos.

@@

We present denoising results on six classic test sequences used commonly in the
literature. We compare against two state-of-the-art methods: V-BM4D
\cite{Maggioni2012}, and the method of Li, Zhang and Dai \cite{LiZhangDai2011}
which we call I-SA-BM4D, since it can be regarded as an iterative version of BM4D \cite{Maggioni2013}
using shape adaptive 3D patches, as in \todo{REF}.  
For V-BM4D, we show results corresponding to two
different choices of parameters. The method labeled V-BM4D-tip
corresponds to the results reported
in \cite{Maggioni2012}. The results labeled V-BM4D-mp were
obtained using the implementation available online \cite{bm4dcode}. This
implementation provides three parameter profiles of increasing quality and
complexity: ``low complexity profile'', ``normal profile'' and ``modified
profile''. We show the results obtained with the modified profile.

Each sequence was contaminated with white Gaussian noise of $\sigma =
10,20,30,40$. Table \ref{tab:psnr-classic-gray} shows the
PSNR values obtained after denoising. For our method, we show results varying the temporal
component of the patch size between $s_t = 1$ (2D patch) and $s_t = 4$ frames.
For our method and V-BM4D-mp we show the PSNR of the basic estimate as well.

@@

\pcomment{TODO: qualitative comparison.}

@@

\subsection{Evaluating temporal consistency}

>>>>

Advances in video sensor hardware have steadily improved the acquisition
quality. However, due to the reduction in the price of cameras
and data storage, video cameras are being used more each time, and in less
favorable situations, such as low lighting conditions. This results in high 
levels of noise, which negatively affects the visual quality of the video and
hinders its use for any application. 
\pcomment{It would be nice to have some reference for these claims about
sensors, camera prices, etc.}

Video denoising has received much less attention than still image denoising in
the literature.  In principle, denoising videos should be easier than still
images, since the strong temporal redundancy along motion trajectories favours
the denoising task. However, this additional form of redundancy creates
challenges.  On one hand, the amount of data is much larger than for still
images, and efficient ways of navigating through this data are needed. On the
other hand, the output of the denoising algorithm should be temporally
consistent according to the motion of the original sequence.

@@

Some methods do not distinguish between the temporal and spatial dimensions,
and treat the video as a volume, which is denoised in a transformed domain
\cite{Rajpoot2003,Wilson2004,Selesnick2003}. 
\pcomment{TODO: Read these references. Even if they consider a 3D transform, I think
they DO distinguish between space and time}

Until the advent of patch-based approaches, methods without motion compensation
failed in the presence of moderate motion, and their main interest resided
in their simplicity and low computational cost. First state-of-the-art results
obtained without motion estimation were reported in \cite{Buades2005v}, with the
non-local means algorithm. The authors argued that for their method, based on
averaging similar patches, motion estimation was not only unnecessary but
counterproductive: while textured regions are problematic for motion
estimation, they are a source of a large number of similar patches across
different frames. Even if the motion trajectory could be reliably estimated, it
makes more sense to use all similar patches, rather than using only the ones on the
trajectory.
%
A similar approach was followed by the authors of \cite{Dabov2007v}. This method is based
on collaborative filtering of similar patches in a
spatio-temporal neighborhood.

The methods in \cite{Buades2005v} and \cite{Dabov2007v} are extensions to video
of still image patch-based denoising algorithms (non-local means \cite{Buades2006} and BM3D
\cite{Dabov2007tip} respectively). They are based on filtering similar 2D patches
searched for in a 3D spatio-temporal neighborhood. These methods exploit both
spatial and temporal redundancy, however, since each 2D patch is processed
independently, there is no mechanism to impose coherence along trajectories.
This results in flickering artifacts which become noticeable for high levels of
noise. 
%
While these methods exploit spatio-temporal redundancy, they do not impose
coherence of the trajectories, resulting in flickering
artifacts which become noticeable for high levels of noise.

@@

In this work we present an extension to video of the non-local Bayes image
denoising method \cite{Lebrun2013a,Lebrun2013ipol} using 3D rectangular
patches and which does not require motion estimation. Yet, it provides
results which compare favourably to the current state-of-the-art,
both quantitatively by a significant marging and qualitatively, specifically in
terms of the temporal consistency attained.
The method works by learning a Gaussian \emph{a priori} model from sets of 
similar noisy patches. Each patch is denoised by computing the \emph{maximum a
posteriori} (MAP) with the estimated prior.
Learning the Gaussian prior amounts to estimating a covariance matrix with a
limited number of samples in a high dimensional space. Motivated by recent 
theoretical results and supported by an empirical analysis, we add the 
assumtion that the \emph{a priori} model is of low rank (\ie the clean patches
lie in a lower dimensional subspace). This assumtion lowers the computational cost
of the algorithm, and improves the results.

@@

The rest of the paper is organized as follows: in \S
\ref{sec:review_nonlocal_Bayes} we review the Bayesian patch denoising principle 
and describe our extension to video. The final algorithm and details about its
implementation are given in \S \ref{sec:algorithm}.
Some relevant aspects of the proposed method are discussed in \S
\ref{sec:discussion}.
In \S \ref{sec:parameters}
we explain the selection of the parameters. Some results, including a
comparison with the state-of-the-art methods, are shown in \S
\ref{sec:results}. Concluding remarks are given in \S \ref{sec:conclusion}.

@@

The Figures \ref{fig:patch_groups_pos} and \ref{fig:patch_groups_patches}
show two examples of groups of similar $9\times 9\times 4$ patches taken from
a sequence contaminated with AGWN of $\sigma = 40$. Figure
\ref{fig:patch_groups_pos} shows the position of 200 nearest neighbors of a
reference patch searched in a $37\times 37\times 5$ search region. The
reference patch is centered in the central frame. The position of the first 5
similar patches are highlighted in red, and they follow the trajectory of the
reference patch throughout the search region.  The next 40 similar patches are
shown in red. In the first example they are spatially shifted versions of the
five nearest neighbors. In the second example, in addition to these shifts
other similar structures in the red sign are detected.  The next nearest
neighbors are shown in blue.

Figure \ref{fig:patch_groups_patches} shows the 20 nearest neighbors in 
more detail. Each patch is displayed by vertically stacking its four $9\times
9$ spatial slices. For the two examples shown in Figure \ref{fig:patch_groups_pos}
the figure shows the noisy patches, the result of the MAP estimation (setting
$r = 40$) and the mean patch together with the leading 19 principal directions.
Since the patch is not motion-compensated, all spatial slices can be different,
and thus the same applies to the principal directions.
The spatial slices of the principal directions show oscillating patterns, with
increasing frequency. In some cases these patterns are localized in a part of
the spatial slice.

@@

\begin{figure*}[htpb!]
	\centering
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}\\
	\vspace{.2cm}
	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
%	\includegraphics[width = .8\textwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_coor.png}
	\caption{Examples of sets of 200 nearest neighbors spatio-temporal patches,
	for patches of size $9\times 9\times 4$.
	The images in each row show the five frames of a $37\times37\times5$ search region. 
	In red we show the positions of the 5 nearest neighbors, in green the next 40 and in blue the rest.
	Note the points shown correspond to the top-left pixel of the first spatial
	slice of each patch.
	The reference patch is the red patch in the center of the search region. To highlight the
	position of the patches, the color of the images has been attenuated.}
	\label{fig:patch_groups_pos}
\end{figure*}

\begin{figure}[htpb!]
	\centering
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_045_085_012_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
	\vspace{.2cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
	\vspace{.2cm}
	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_255_056_010_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
	\vspace{.2cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_deno.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_nisy.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_orig.png}\\
%	\vspace{.1cm}
%	\includegraphics[width = \columnwidth]{figs/patch_groups/patch_group_bus_110_090_010_s40_wx37_wt2_sx9_st4_r040_n200_pcas.png}\\
%	\vspace{.1cm}
	\caption{For each of the two sets of patches of Fig. \ref{fig:patch_groups_pos} we show
	the 20 nearest neighbors (top), the corresponding MAP estimates (middle) and the sample mean 
	and first 19 principal directions computed from the sample covariance matrix (bottom).
	For each image, each column corresponds to the 4 spatial slices of a $9\times9\times4$ patch.}
	\label{fig:patch_groups_patches}
\end{figure}

\section{Description of the algorithm}
\label{sec:algorithm}

\subsection{Implementation details}
\label{sse:implementation}

\subsection{Computational complexity and speed-ups}
\label{sse:complexity}

Let us consider the complexity of constructing and filtering a group of
$n$ similar patches.
Processing each group of similar patches is of asymptotic order $\mathcal
O(w_x^2 w_t s_x^2 s_t + ns_x^4s_t^2 + s_x^4s_t^2r)$.
The first term comes from the exhaustive search of the $n$ nearest patches in
the search window.  The remaining terms correspond respectively to the
computation of the covariance matrix and its truncated eigen-decomposition.

The most expensive operations are the computation of the covariance matrix and
of its principal directions.
Let us note that there are methods for computing the $r$ principal directions
of a set of points that do not require computing the covariance matrix
\cite{Halko2011}. Instead these methods efficiently approximate a truncated SVD
of the data matrix, and would 
allow to improve the computational complexity to $\mathcal O(w_x^2 w_t s_x^2
s_t + s_x^2s_trn)$.

The computation can be substantially accelerated by reducing the 
number of groups of patches that are jointly processed.
Each group of patches is associated to the reference patch from which 
the nearest neighbors are computed.
We apply two tricks
considered in \cite{Lebrun2013ipol} in order to reduce the number of reference patches. 
The first one consists on visiting only 
those patches on a coarser spatial grid with vertical and horizontal separation of $s_x/2$
(similar to \cite{Dabov2007tip}). This reduces the number of reference patches by a
factor of $s_x^2/4$. The second trick consists on the following.  After
filtering the patches in a group, these patches are not considered again
as reference patches. 
Note that we ensure that each reference patch is filtered at least once. 
Due to the spacing of the coarse grid, this guarantees that each pixel is
overlapped by four reference patches on the same frame, so it will receive at
least four contributions in the aggregation. The actual number of aggregated
estimates for each pixel is much higher, considering the reference and
non-reference patches (which can also be filtered) in the neighboring frames.

Due to these speed-ups, the actual number of groups of similar patches processed
depends on $n$, and varies between a $1\%$ ($n \approx 400$)  and a $10\%$ ($n
\approx 50$) of the total number of pixels in the sequence. 
\pcomment{Check this numbers!}

\section{Discussion}
\label{sec:discussion}

\subsection{Spatio-temporal patches}
\label{sse:spatio-temporal-patches}
Patch-based methods filter similar patches to remove noise. When the noise is
Gaussian, the patch similarity is determined based on the (squared) Euclidean distance 
between patches. Due to the noise, the Euclidean distance follows a non-central
$\chi^2$ distribution, with variance
\[\textnormal{Var}\left\{\frac1{m}\|\ma q_1 - \ma q_2\|^2\right\} = \frac{8\sigma^2}{m}\left(\sigma^2 + \frac1{m}\|\ma p_1 - \ma p_2\|^2\right),\]
where $\ma q_1, \ma q_2$ are the noisy versions of the patches $\ma p_1, \ma p_2$, 
and $m$ denotes the number of pixels in the patch. The noise in the patch comparison
can be reduced by considering larger patches.
However, increasing the size of the patch also increases the distances between
patches and reduces the likelihood of finding similar patches.
%
The additional temporal dimension in a spatio-temporal patch allows to increase
the number of pixels in the patch, without increasing its spatial size.
Due to the high redundancy of the video in the temporal dimension, increasing the 
temporal size of the patch causes a much lower increase in the patch distances.
%The high redundancy of the video in the temporal dimension, allows to have 
%larger patches without affecting much the number of nearest neighbors of the patch.
%

@@

% \pcomment{
% \begin{itemize}
% %	\item Justify why 3D patches: Greater distance noise reduction with roughly the same 
% %		degree of self-similarity. One way to see this is that the 3D patches can be 
% %		seen, for a video in which the brightness constancy assumption holds, as a 2D patch
% %		by back-projecting all the slices to the first frame.
% 	\item Relation with respect other iterative methods found in the literature
% 		(SOS, twicing, diffusion, BM3D, I-SA-BM4D).
% %	\item Variation of the results tables for different sequences.
% \end{itemize}}

\subsection{Estimation of the covariance matrix}
\label{sse:covariance-estimation}

Learning the parameters of the \emph{a priori} distribution for each group of
noisy patches amounts to estimating a covariance matrix is a high dimensional space 
and with a limited number of samples. 
%The latter is controlled by the $n$ parameter, which
%should not be too high in order for the Gaussian prior to be valid.
%
The empirical covariance matrix $\widehat C$ converges to the true covariance matrix
$C$ as the number of samples from the underlying Gaussian distribution grows to infinity.
But in our case, the number of samples we can ``draw'' from the Gaussian is limited:
%The Gaussian is only an approximate density model for the set of
%$n$ nearest neighbors of a target patch.
As we increase the number $n$ of neighbors of a patch, we include patches
that are not as similar to the reference patch and the Gaussian model becomes less valid.

Thus there is a trade-off between accuracy of Gaussian model and precision of the
covariance matrix. We will find a compromise to this trade-off empirically  in
\S \ref{sec:parameters}. But it still intersting to review the literature 
on the sample covariance matrix as an estimator of the population covariance.

\pcomment{Results of Bunea are still missing!}
\pcomment{More detailed explanation of the results?}

\section{Parameter selection}
\label{sec:parameters}

@@

\subsection{Search region and patch size}

Setting the size of the search region implies a
trade-off between denoising performance and running time: A larger search
region takes longer to traverse but yields better results, since more
similar patches can be found. We fix the search
region for both stages of the algorithm at 
\begin{equation}
w_x = 37 \quad \text{ and } \quad w_t = 5.
\label{eq:search_region_parameters}
\end{equation}
With these values, the search region captures the $5$ frame trajectory of the
target patch as long as it moves slower than $37/5 = 7.4$
pixels/frame.

Let us now consider the spatial patch
size $s_x$, which is a key parameter in patch-based methods \cite{Buades2006,Kervrann2006,Boulanger2007}.
%The distribution of the 
%patch distances is a non-central Chi-squared distribution \cite{Dabov2007tip}
%\pcomment{CHECK!}
%with variance 
%\[\textnormal{Var}\left\{\frac1{m}\|\ma q_1 - \ma q_2\|^2\right\} = \frac{8\sigma^2}{m}\left(\sigma^2 + \frac{\|\ma p_1 - \ma p_2\|^2}{m}\right),\]
%where $m = s_x^2s_t$ is the total number of pixels in the patch.
Increasing the spatial patch size $s_x$ reduces the variance of the noise in the patch
distance. 
However, for larger
patches it becomes less likely to find other patches with a similar texture
pattern. The use spatio-temporal patches helps in solving this trade-off: for
instance, with a spatio-temporal patch of size $5\times 5\times 4$, we obtain
the same patch distance noise as with a spatial patch of size $10\times 10
\times 1$. Moreover, as discussed in \S \ref{sse:spatio-temporal-patches}, for
an approximately uniform motion, the 
amount of similar neighbors of a $5\times 5\times 4$ patch is roughly the same
as for a $5\times 5\times 1$ patch.

@@

\subsection{Rank and number of similar patches}

The number of similar patches $n$ and the rank $r$ affect the estimation of the covariance
matrix $C$, controlling the trade-off between the accuracy of Gaussian model
and precision of the covariance matrix,
which in turn has a direct effect on the denoising strength of the algorithm.
% This have great impact on the bias-variance trade-off of the denoiser.
%In practice, these parameters determine the denoising strength of the
%algorithm. For example, using a large
We find a compromise to this trade-off empirically, by trying different values of this parameter on a 
few training sequences and use the one yielding the best PSNR.

%The number of similar patches $n$ and the rank $r$ affect the estimation of the covariance
%matrix $C$, controlling the trade-off between the accuracy of Gaussian model
%and precision of the covariance matrix.

@@

The plots in Figure \ref{fig:rank1-nsim1-bpsnr} show the PSNR of the basic
estimate as a function of the rank $r_1$, for several values of $n_{1}$.
The curves correspond to an average over 20 frames clips 
extracted from four classical test
sequences, namely \emph{bus}, \emph{foreman}, \emph{football} and \emph{tennis}.
The sequences were contaminated with AGWN of standard deviation $\sigma = 10,
20, 30, 40$. 
%
%
We can interpret the results in the light of the theoretical results
described in \S \ref{sse:covariance-estimation}.
For all noise levels, best results are obtained when using a larger number of
samples $n_1$. All curves of fixed $n_1$ show a maximum at a given rank lower than
the dimensionality of the patch. As the noise level increases, this optimal rank
shifts to the left. This makes sense, since less eigenvectors are reliable estimates.
%
Also, for a fixed noise level, the optimal rank is smaller for smaller values of $n_1$.
Smaller groups obtained by keeping only the
most similar patches have less modes of variation, and thus a worse signal-to-noise ratio
in the eigenvectors with lower eigenvalues.
%As the group of patches grows
%bigger, patches in the group are not as similar anymore, and a higher rank is
%needed to allow for additional modes of variation.

@@

In this section we present results obtained with the proposed
method on some classical test sequences with Gaussian noise added.
We consider separatedly the cases of grayscale and color videos.

@@

We present denoising results on six classic test sequences used commonly in the
literature. We compare against two state-of-the-art methods: V-BM4D
\cite{Maggioni2012}, and the method of Li, Zhang and Dai \cite{LiZhangDai2011}
which we call I-SA-BM4D, since it can be regarded as an iterative version of BM4D \cite{Maggioni2013}
using shape adaptive 3D patches, as in \cite{Ercole2005}.  
For V-BM4D, we show results corresponding to two
different choices of parameters. The method labeled V-BM4D-tip
corresponds to the results reported
in \cite{Maggioni2012}. The results labeled V-BM4D-mp were
obtained using the implementation available online \cite{bm4dcode}. This
implementation provides three parameter profiles of increasing quality and
complexity: ``low complexity profile'', ``normal profile'' and ``modified
profile''. We show the results obtained with the modified profile.

Each sequence was contaminated with AGWN of $\sigma =
10,20,30,40$. Table \ref{tab:psnr-classic-gray} shows the
PSNR values obtained after denoising. For our method, we show results varying the temporal
component of the patch size between $s_t = 1$ (2D patch) and $s_t = 4$ frames.
For our method and V-BM4D-mp we show the PSNR of the basic estimate as well.

@@

For a qualitative comparison, Figure \ref{fig:results_mobile_mono}
shows two details of the results obtained by our method and V-BM4D for the
\emph{mobile} sequence, for noise of $\sigma = 10, 20, 40$. It can be observed
that the proposed method preserves more details. It also exhibits a higher
temporal consistency, as will be shown in \S \ref{sse:temporal-consistency}.

@@

\subsection{Evaluating temporal consistency}
\label{sse:temporal-consistency}
